Graph Topological Aspects of Granger Causal Network Learning

### 2019-09-29 presentations/mcgillOct2019/pres.tex
More figures for presentation.

\documentclass{beamer} %With pauses
% \input{\string~/Documents/academics/global_academics/latex_preamble}
%\documentclass[handout]{beamer} %Without pauses
\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cancel}
\usepackage{centernot}
\usefonttheme[onlymath]{serif}

@@

  \begin{proposition}
    \label{prop:pwgc_anc}
    If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
  \end{proposition}\pause

@@

  A heuristic ``definition'' (to avoid introducing more cumbersome notation):
  \begin{definition}[Persistent]
    We will say that the process $x(t)$ with Granger causality graph
    $\gcg$ is \textit{persistent} if every node has a self-loop,
    i.e. $\B_{ii}(z) \ne 0$.
  \end{definition}

\end{frame}

\section{Pairwise Testing Recovery Methods}
\begin{frame}
\end{frame}

>>>>

\documentclass{beamer} %With pauses
% \input{\string~/Documents/academics/global_academics/latex_preamble}
%\documentclass[handout]{beamer} %Without pauses
\usepackage[backend=bibtex]{biblatex}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cancel}
\usepackage{centernot}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{subcaption}

\usefonttheme[onlymath]{serif}

@@

  \begin{proposition}[Partial Sufficient Conditions]
    \label{prop:pwgc_anc}
    If $\gcg$ is a strongly causal DAG then $j \in \anc{i} \Rightarrow j \pwgc i$.
  \end{proposition}\pause

@@

  A heuristic ``definition'' (to avoid introducing more cumbersome notation):
  \begin{definition}[Persistent]
    We will say that the process $x(t)$ with Granger causality graph
    $\gcg$ is \textit{persistent} if every node has a self-loop,
    i.e. $\B_{ii}(z) \ne 0$.
  \end{definition}
\end{frame}

\begin{frame}{Persistent Systems}
  Define $H_i(z)$ to be the filter s.t.:

  \begin{equation*}
    H_i(z)x_i(t) = \linE{x_i(t)}{\H_{t - 1}^{(i)}},
  \end{equation*}\pause

  And then with $\sigma_k^2 = \E v_k(t)^2$
  \begin{equation*}
    \label{eqn:T_filter}
    T_{ij}(z) = \sum_{k \in \anc{i} \cap \anc{j}}\sigma_k^2\A_{ik}(z^{-1})(1 - H_i(z^{-1}))(1 - H_j(z))\A_{jk}(z)
  \end{equation*}\pause

  \begin{proposition}
    \label{prop:persistence_converse}
    Fix $i, j \in [n]$ and suppose $\exists k \in \anc{i} \cap \anc{j}$
    which confounds $i, j$.  Then, if $T_{ij}(z)$ is not causal we have
    $j \pwgc i$, and if $T_{ij}(z)$ is not anti-causal we have
    $i \pwgc j$.  Moreover, if $T_{ij}(z)$ is two-sided, then $j \pwgc i \iff i \pwgc j$.
  \end{proposition}
\end{frame}

\section{Pairwise Testing Recovery Methods}
\begin{frame}{Pairwise Recovery}
  \begin{itemize}
    \item{Persistant systems where $T_{ij}(z)$ is not two-sided are highly pathological}\pause
    \item{However, that the following requires an assumption on
        $T_{ij}(z)$ is an unfortunate blemish on the theory as it is
        not a topological requirement}\pause
  \end{itemize}

    \begin{theorem}[Pairwise Recovery]
      \label{thm:scg_recovery}
      If the Granger causality graph $\gcg$ for the process $x(t)$ is
      a strongly causal DAG and $T_{ij}(z)$ is two-sided for every
      $i, j$, then $\gcg$ can be recovered from pairwise causality
      tests alone.
    \end{theorem}

    \begin{itemize}
      \item{Proof is constructive in that I directly provide an algorithm to do so.}
    \end{itemize}
\end{frame}

\begin{frame}{Pairwise Recovery}
  \begin{algorithm}[H]
    \SetKwInOut{Input}{input}
    \SetKwInOut{Output}{output}
    \SetKwInOut{Initialize}{initialize}
    \DontPrintSemicolon

    \footnotesize
    % \BlankLine
    % \caption{Pairwise Granger Causality Algorithm}
    \label{alg:pwgr}
    % \TitleOfAlgo{Pairwise Graph Recovery}
    % \Input{Pairwise Granger causality relations between a persistent
    % process of dimension $n$ whose joint Granger causality
    % relations are known to form a strongly causal DAG $\gcg$.}
    % \Input{Pairwise Granger causality relations}
    % \Output{Edges $\gcge = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of
    %   the graph $\gcg$.}
    \Initialize{$S_0 = [n]$, $E_0 = \emptyset$, $k = 1$}\pause
    \BlankLine
    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$\\\pause
    $P_0 \leftarrow \{i \in S_0\ |\ \forall s \in S_0\ (s, i) \not\in W\}$\\\pause
    \While{$S_{k - 1} \ne \emptyset$}{
      $S_k \leftarrow S_{k - 1} \setminus P_{k - 1}$\\
      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$\\\pause

      $D_{k0} \leftarrow \emptyset$\\\pause
      \For{$r = 1, \ldots, k$} 
      {
        $Q \leftarrow E_{k - 1} \cup \big(\bigcup_{\ell = 0}^{r - 1} D_{k\ell}\big)$\\\pause
        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times P_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } Q\}$\\\pause
      }
      $E_k \leftarrow E_{k - 1} \cup \big(\bigcup_{r = 1}^k D_{kr}\big)$\\\pause
      $k \leftarrow k + 1$\\\pause
    }
    \Return{$E_{k - 1}$}
  \end{algorithm}
\end{frame}

\begin{frame}{Pairwise Recovery}
  An Example (TODO: Intersperse Pauses)

  \begin{columns}
    \begin{column}{0.5\linewidth}
      \begin{figure}
        \includegraphics[width=\linewidth]{../../figures/example_algorithm.pdf}
      \end{figure}
    \end{column}

    \begin{column}{0.5\linewidth}
      \begin{figure}
        \includegraphics[width=\linewidth]{../../figures/example_algorithm2.pdf}
      \end{figure}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Simulation}
  \begin{itemize}
    \item{Construct random graphs and compare PWGC against LASSO method}\pause
    \item{I obtained the best results with Zou's Adaptive LASSO, but known to suffer somewhat from false positives}
    \item{\fullcite{adaptive_lasso_zou2006}}
  \end{itemize}
\end{frame}

\begin{frame}{Simulation}
  Representative Random Graph Topologies on $n = 50$ Nodes.  Parameter
  $q$ controls the density of edges.
  \begin{columns}
    \begin{column}{0.33\linewidth}
      \begin{figure}
        \caption{SCG}
        \includegraphics[width=\linewidth]{../../figures/example_scg.pdf}
      \end{figure}
    \end{column}\pause
    \begin{column}{0.33\linewidth}
      \begin{figure}
        \caption{DAG $(q = \frac{2}{n})$}
        \includegraphics[width=\linewidth]{../../figures/example_dag.pdf}
      \end{figure}
    \end{column}\pause
    \begin{column}{0.33\linewidth}
      \begin{figure}
        \caption{DAG $(q = \frac{4}{n})$}
        \includegraphics[width=\linewidth]{../../figures/example_dag_q.pdf}
      \end{figure}
    \end{column}\pause
  \end{columns}
\end{frame}

\begin{frame}{Simulation}
  $\VAR(p)$ model with $n = 50, p = 15$ and randomized edge placements.  Left: SCG, Right: DAG ($q = 2/n$)
  \begin{columns}
    \begin{column}{0.5\linewidth}
      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_scg_pmax15_simulation_mcc_only.png}
    \end{column}

    \begin{column}{0.5\linewidth}
      \includegraphics[width=0.9\linewidth]{./figures/new_lasso_comparison_dag_pmax15_simulation_mcc_only.png}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Simulation}
  \includegraphics[width=\linewidth]{../../figures/new_increasing_n_simulation.pdf}\pause
\end{frame}

\begin{frame}{Simulation}
  \includegraphics[width=\linewidth]{../../figures/new_mcc_comparison001.pdf}
\end{frame}
